Whats inside the news?

### 2017-12-11 phd_slides/phd-workshop.Rmd
slides

```{r include=FALSE}
suppressPackageStartupMessages({
  library(dplyr)       # Data manipulation
  library(stringr)     # String manipulation
  library(lubridate)   # Date and time manipulation
  library(purrr)       # Functional programming
  library(tidyr)       # Reshaping
  library(magrittr)    # Advanced piping
  library(pushoverr)   # Pushover notifications
  library(doMC)        # Parallel Computing
  library(readr)       # Importing data
  library(tibble)      # Better data frames
  library(data.table)
  library(pscl)
  library(boot)
  library(stm)
  library(tm)
  library(readxl)

@@

  library(tidytext)    # Tidy text mining
  library(stringdist)  # String distances
  library(proxy)       # Distance measures
})

@@

# Introduction
## Online News
<div style="position:relative; width:640px; height:480px; margin:0 auto;">
  <img class="fragment" src="../figs/screenshot-spon.png" style="position:absolute;top:0;left:0;" />
  <img class="fragment" src="../figs/screenshot-focus.png" style="position:absolute;top:0;left:0;" />
  <img class="fragment" src="../figs/screenshot-bild.png" style="position:absolute;top:0;left:0;" />
  <img class="fragment" src="../figs/screenshot-tagesschau.png" style="position:absolute;top:0;left:0;" />
</div>

@@

<img class="fragment" src="../figs/onlinenews-2.png" width="900"/>

@@

<img class="fragment" src="../figs/timeline.png" width="850"/>

----------------------
```{r eval=FALSE, include=FALSE}
load("../output/models/STM 28 .Rda")

tidy.df <- btw %>%
  unnest_tokens(word, text_cleaned) %>%
  count(word, articleID, sort =TRUE)

dtm <- tidy.df %>% dplyr::filter(nchar(word)>1) %>%
  cast_dtm(articleID, word, n)
save(stmOut, btw, dtm, tidy.df, file="../output/presentation_files.Rda")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
load("../output/presentation_files.Rda")
table <- inspect(dtm[1:5,1:10])
```

> - Documents (articles) are stored on a "Document-Term-Matrix"
```{r echo=FALSE}
htmlTable::htmlTable(table)
```

@@

- How to find out latent topics in an article?

@@

<img src="../figs/lda-2.png">
<small>Credits: Christine Doig</small>

@@

<img src="../figs/lda-blei.png" width="750"/>

@@

<aside class="notes">
 - Documents are seen as "bag of words". 
 - Mixed memberhip model: Population of topics stays the same, but distrubution over topics changes for ech document
 - Each document is comming from a mixture model, where the mixture proportions change from document to document but mixture components are fixed a cross all documents.
 - Each document arises as follows: 
    1. chose a distribution over topics (histogram) 
    2. For each word in the document, chose a colored button
    3. look up the distribution over terms associated with that button and draw a word
</aside>

@@

<img src="../figs/lda-plate.png" width="800">

> - Nodes a random variables; edges indicate dependence
> - Shaded nodes are observed; unshaded nodes are hidden.
> - Plates indicate replicated variables.

<aside class="notes">
		-Generative Process:
  1. chose Number of topics (e.g. K=100)
  2. Each topic has a $\beta$, which is a distribution over the full vocabulary
  3. For each Document (D) I chose a distribution over topics ($\theta$)
  4. For each word in each document (N) I chose a colored button from $\theta$ 
  5. Chose a word from the corresponding distribution over terms
- Diagram:
    - defines a factorization of the joint probability distribution of hidden and observed random variables.
    - encodes indepence assumptions about the variables (which variables are dependent and conditionally independent) 
    - Connects to algorithm for computing with data: Inference problem: finding the hidden variables given the observations.
	</aside>

------------------------
<img src="../figs/lda-plate-2.png" width="800">

- The posterior is: $p(\theta,z,\beta|w) = \frac{p(\theta,\beta,z,w)}{p(w)}$

    - **numerator**: joint distribution of all random variables (can be easily computed for any setting of hidden variables.)
    - **denominator**: marginal probability of the observations (probability of seeing the observed corpus under any topic model.)

- From a collection of documents, infer
    - Per-word topic assignment $z_{d,n}$
    - Per-document topic proportions $\theta_d$
    - Per-corpus topic distributions $\beta_k$

@@

<aside class="notes">
  - the joint distribution defines a posterior. 
  - In theory, it can be computed by summing the joint distribution over every possible instantiation of the hidden topic structure.
  - $p(\theta,z,\beta|w)$ the probability of the hidden structures given the observed words in a document.
  - - That number of possible topic structures, however, is exponentially large; this sum is intractable to compute. As for many modern probabilistic models of interest—and for much of modern Bayesian statistics—we cannot compute the posterior because of the denominator, which is known as the evidence. A central research goal of modern probabilistic modeling is to develop efficient methods for approximating it. Topic modeling algorithms—like the algorithms used to create Figures 1 and 3—are often adaptations of general-purpose methods for approximating the posterior distribution.
  - First uncover the structure (calculate th posterior) - use the discovered distributions to perform task
</aside>

@@

<img class="fragment" src="../figs/research_strategy.png" width="1000"/>

@@

```{r Extract wtp and dtp, include=FALSE}
load("../output/models/STM 28 .Rda")

# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp
# Document-topic probabilities
stmOut %>% tidy("gamma") -> posts.dtp

@@

<img src="../figs/topic_proportion.png" width="600">

>>>>

```{r include=FALSE}
suppressPackageStartupMessages({
  library(dplyr)       # Data manipulation
  library(stringr)     # String manipulation
  library(lubridate)   # Date and time manipulation
  library(magrittr)    # Advanced piping
  library(pushoverr)   # Pushover notifications
  library(readr)       # Importing data
  library(data.table)
  library(pscl)
  library(boot)
  library(stm)
  library(tm)
  library(readxl)
  library(tidytext)
  library(MCMCpack)
  library(Compositional)

@@

  library(stringdist)  # String distances
  library(proxy)       # Distance measures
})

@@

# Introduction
## Online News

<div style="position:relative; width:640px; height:480px; margin:0 auto;">
  <img class="fragment" src="img/screenshot-spon.png" style="position:absolute;top:0;left:0;" />
  <img class="fragment" src="img/screenshot-focus.png" style="position:absolute;top:0;left:0;" />
  <img class="fragment" src="img/screenshot-bild.png" style="position:absolute;top:0;left:0;" />
  <img class="fragment" src="img/screenshot-tagesschau.png" style="position:absolute;top:0;left:0;" />
</div>

@@

<img class="fragment" src="img/onlinenews-2.png" width="900"/>

@@

<img class="fragment" src="img/timeline.png" width="850"/>

## Concepts

- A single observation in a textual database is called a *document*.

- The set of documents that make up the dataset is called a *corpus*.

- Covariates associated with each document are called *metadata*.

## Data Structure
```{r message=FALSE, warning=FALSE, include=FALSE}
load("data/presentation-data.Rda")
table <- inspect(dtm[1:5,1:10])
```

- Documents (articles) are stored on a "Document-Term-Matrix"
-
```{r echo=FALSE}
htmlTable::htmlTable(table)
```

@@

## How to find out latent topics in an article?

@@

<img src="img/lda-2.png">
<small>Credits: Christine Doig</small>

@@

<img src="img/lda-blei.png" width="750"/>

@@

<aside class="notes">
Mixed memberhip model: Population of topics stays the same, but distrubution over topics changes for each document.
Each document is comming from a mixture model, where the mixture proportions change from document to document but mixture components are fixed a cross all documents.
</aside>

@@

* Nodes are random variables; arrows indicate dependence

* Plates indicate replicated variables:
    * $N =$ collection of words within a document.
    * $D =$ collection of documents within a corpus.

* Shaded nodes are observed; unshaded nodes are hidden
    * observed: word in a document $w_{d,n}$
    * fixed: mixture components (number of topics $K$ & vocabulary)
    * hidden: mixture proportions (per-document topic proportions $\theta_d$ & word-topic distribution $\beta_k$)

----------------------------
<img src="img/lda-plate-2.png" width="600">
----------------------------

<aside class="notes">
Diagram:
(1)defines a factorization of the joint probability distribution of hidden and observed random variables.
(2)encodes indepence assumptions about the variables (which variables are dependent and conditionally independent) 
(3)Connects to algorithm for computing with data: Inference problem: finding the hidden variables given the observations.
</aside>

## Generative Process I
* Generative Process for the $n^{th}$ word-position in document $d$: 
    1. $K$: choose the number of topics;
    2. $\theta_d$: for each document $d$, choose a distribution over topics;
    3. $z_{d,n}$: according to $\theta_d$, assign a topic for the $n^{th}$ word;
    4. $w_{d,n}$: choose a term from that topic according to $\beta_k$
    5. $N$: repead this process for all $n$ word-positions in the document.
    6. $D$: conduct this process for all $d$ documents in the corpus

## R Markdown

test 1

***

test 2

## Generative Process II

$p(\beta_{1:K},\theta_{1:D},z_{1:D}, w_{1:D})=\displaystyle \prod_{i=1}^{K}p(\beta_i)\displaystyle \prod_{d=1}^{D}p(\theta_d)(\prod_{n=1}^Np(z_{d,n,}|\theta_d)p(w_{d,n}|\beta_{1:K},z_{d,n}))$

```{r}
draws <- rdirichlet(200, c(1,1,1) )
bivt.contour(draws)
```

## Bayesian inference

Posterior probability: $P(A|B)=\frac{P(B|A)P(A)}{P(B)}$

$p(\theta_{1:D},z_{1:D},\beta_{1:K}|w_{1:D}) = \frac{p(\theta_{1:D},\beta_{1:K},z_{1:D})(w_{1:D})}{p(w_{1:D})}$

  * **numerator**: joint distribution of all random variables (can be easily computed for any setting of hidden variables.)
  * **denominator**: marginal probability of the observations (probability of seeing the observed corpus under any topic model.)

@@

<aside class="notes">
the joint distribution defines a posterior. 
In theory, it can be computed by summing the joint distribution over every possible instantiation of the hidden topic structure.
$p(\theta,z,\beta|w)$ the probability of the hidden structures given the observed words in a document.
That number of possible topic structures, however, is exponentially large; this sum is intractable to compute. As for many modern probabilistic models of interest—and for much of modern Bayesian statistics—we cannot compute the posterior because of the denominator, which is known as the evidence. A central research goal of modern probabilistic modeling is to develop efficient methods for approximating it. Topic modeling algorithms—like the algorithms used to create Figures 1 and 3—are often adaptations of general-purpose methods for approximating the posterior distribution.
First uncover the structure (calculate th posterior) - use the discovered distributions to perform task
</aside>

@@

<img class="fragment" src="img/research_strategy.png" width="1000"/>

@@

```{r Extract wtp and dtp, include=FALSE}
# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp
# Document-topic probabilities
stmOut %>% tidy("gamma") -> posts.dtp

@@

<img src="img/topic_proportion.png" width="600">
